Analysis of Cartesian Genetic Programming's Evolutionary Mechanisms

### 2013-05-22 CGP Journal/main.tex
Wrote the CGP introduction section, added some citations

\section{Cartesian Genetic Programming}
TODO Explain basic CGP, use~\cite{miller:2011:chapter2}.

@@

>>>>

\section{Cartesian Genetic Programming}
Cartesian Genetic Programming (CGP) was originally proposed as a method for general
genetic programming in~\cite{miller:2000:CGPorigin}.  While its beginnings and
its name come from evolving circuits on a two dimensional grid, modern CGP can
represent any directed acyclic graph (DAG), with applications in evolving binary
circuits~\cite{walker:2008:cgpmodules},
robot controllers~\cite{harding:2005:robots},
neural networks~\cite{khan:2010:cgpann},
image classifiers~\cite{harding:2012:mtcgp},
and regression~\cite{harding:2009:smcgp}.

CGP represents DAGs using a linear genome of integer values.  Each node in the
DAG is encoded using a collection of genes, with one gene specifying the function
the node applies to its inputs, and the remaining genes expressing where the
node takes input from.  Nodes can take input from any problem input and from
any node preceding them in the genome.  To complete the representation, a set
of extra genes are included at the end of the genome to specify which nodes or
input locations to use as function outputs.

Unlike tree based GP encodings, CGP is able to perform value reuse.  This means
that the output of a node can be reused by any number of other nodes as input values.
CGP's requirement that nodes connect forward in the genome prevents this ability
from causing cycles.

As output locations and information flow in the DAG are evolvable, there are often
large portions of the genome which do not participate in creating any of the
output values.  These nodes are referred to as inactive, with the nodes currently
being used to create output values referred to as active.  Inactive nodes
allow CGP individuals to drift genetically, as they can be mutated without effecting
the fitness of the individual.  This genetic drift can then be incorporated, as
mutation can change the DAG structure causing previously inactive nodes to become
active.  Previous work suggests CGP is most efficient when up to 95\% of the genome
is inactive~\cite{miller:2006:redundancy}, yet other work suggests this may
be a result of hidden parsimony pressure in CGP~\cite{goldman:2013:ordering}.

CGP uses very simple evolutionary mechanisms.  By far the most common strategy
is $1+4$, in which each generation a single parent produces four offspring using
mutation.  The best offspring then competes with the parent, with the offspring
replacing the parent if it is no less fit.  This replacement strategy encourages
drift, as completely neutral mutations are always preferred to stagnation.  Mutation
is commonly performed by mutating each gene at a set probability, where mutation
involves changing a gene randomly to some different valid value.  For instance,
if a function gene is chosen for mutation, its new value is randomly chosen from
all possible functions, with the exception of the genes current value.  Combined,
this form of CGP can be viewed as a stochastic hill climber with neutrality.
For a more in depth description of CGP, see~\cite{miller:2011:chapter2}.

@@

TODO Include scaling comparison with~\cite{walker:2008:cgpmodules}.
